---
title: "Homework2 Code"
output: html_document
---

#Question 1
lambda=15 (mean)
To set up the distribution, does it matter what "n" is? Should I just pick a big enough number to get a good distribution?

set.seed(1)
x <- rpois(100000, 15)
plot(x, dpois(x, 15, log=FALSE))

Okay so probability that she'll hear more than 8 calls...

ppois(x>=8, 15, lower.tail=TRUE, log.p=FALSE)
I could do a sum but there's no upper limit

or could I use qpois?

qpois(p, lambda, lower.tail=TRUE, log.p=FALSE)

qpois(0.975, 15, lower.tail=TRUE, log.p=FALSE)
=23...so can I asssume that 23 is a good upper bound, because 95% of the values will be under 23? if so:

sum(ppois(8:23, 15, lower.tail=TRUE, log.p=FALSE))

Unsure, I got a number larger than 1, not sure if that can happen.

According to the textbook:
```{r echo=TRUE}
1-ppois(q=8, lambda=15)
```


BLESS MY SOUL I think I got it :')

#Question 2
Probability of hearing no calls
Before, I tried: p(x)= lambda^x*exp(-lambda)/x! but I got a number that was x10^-7

What if I do ppois again?

ppois(0, 15, lower.tail=TRUE, log.p=FALSE)

Update I got the same number as doing p(0) so maybe that's the right answer...

According to  the textbook:
```{r echo=TRUE}
dpois(x=0,lambda=15)
```



#Question 3
exactly 3 calls in a session, x=3, I know I can use that formula.

p3 <- 15^3*exp(-15)/factorial(3)
p3

According to the textbook:
```{r echo=TRUE}
dpois(x=3,lambda=15) #output is the same as the above formula
```


ppois(3, 15, lower.tail=TRUE, log.p=FALSE)
This one was about 2x larger than the p(3) formula. BECAUSE it's left cumulative, so adding up probabilities of 0-3

#Question 4
plot values in the range 0:30

y<- seq(0, 30, 0.0001)
plot(y, dpois(y, 15, log=FALSE), cex=0.4)

Okay so now the x axis goes from 0-30 which is what I wanted, but there's a black line near the bottom, not sure why it's there, and it looks slightly like a normal distribution
Also I'm getting warnings and I don't know why/what that means...update, some of them are non-integers? Why would that be if log=FALSE?

According to the textbook:
```{r echo=TRUE}
plot(dpois(x=0:30,lambda=15),ylim=c(0,0.25),names.arg=0:30, ylab="Pr(X=x)", xlab="x")
```


Okay so it's saying that I have non-graphical parameters but I'm getting a graph?

#Question 5
simulate 104 results (2 years worth of work)
so would n=104 or 1? Because you want one result per sample (not multiple numbers from the vector per sampling?)

Would the formula/code be the same for different types of distributions?
```{r echo=TRUE}
set <- NULL
n<-1
for (i in 1:104){
set[i] <- mean(sample(x, n, replace=TRUE))
}
quantile(set)
quantile(set, c(0.025, 0.975))
```

I think I did it? Maybe without the quantile parts?

#Question 6
Honestly I'm totally winging it at this point

sooooo creating a histogram of the simulated results
```{r  echo=TRUE}
hist(set, xlim=c(0,30))
```
