---
title: "Module11 Notes"
author: "Dani Antos"
date: "October 7, 2017"
output: html_document
---

```{r}
library(curl)
library(ggplot2)
library(manipulate)

```

Recap:

-Type I error: falsely reject the null; the probability of this error is equal to alpha (the significance level)

-we can decrease the chance of this   error by decreasing alpha

-Type II error: falsely accept the null, probability of this error is called beta, and we usually don't have a specific value for it

#Type I Error
typeI() will evaluate the Type I error rate
```{r Type 1 error function}
typeI <- function(mu0, sigma, n, alternative = "two.tailed", alpha = 0.05, k = 10000) {
}
typeI <- function(mu0, sigma, n, alternative = "two.tailed", alpha = 0.05, k = 1000) {
    p <- rep(NA, k)  # sets up a vector of empty p values
    for (i in 1:k) {
        # sets up a loop to run k simulations
        x <- rnorm(n = n, mean = mu0, sd = sigma)  # draws a sample from our distribution
        m <- mean(x)  # calculates the mean
        s <- sd(x)  # calculates the standard deviation
        z <- (m - mu0)/(s/sqrt(n))  # calculates the T statistic for the sample drawn from the null distribution relative to the null distribution
        # alternatively use t <- (m-mu0)/(s/sqrt(n))
        if (alternative == "less") {
            p[i] <- pnorm(z, lower.tail = TRUE)  # calculates the associated p value
            # alternatively, use p[i] <- pt(t,df=n-1,lower.tail=TRUE)
        }
        if (alternative == "greater") {
            p[i] <- pnorm(z, lower.tail = FALSE)  # calculates the associated p value
            # alternatively, use p[i] <- pt(t,df=n-1,lower.tail=FALSE)
        }
        if (alternative == "two.tailed") {
            if (z > 0) 
                {
                  p[i] <- 2 * pnorm(z, lower.tail = FALSE)
                }  # alternatively, use if (t > 0) {p[i] <- pt(t,df=n-1,lower.tail=FALSE)}
            if (z < 0) 
                {
                  p[i] <- 2 * pnorm(z, lower.tail = TRUE)
                }  # alternatively, use if (t < 0) {p[i] <- pt(t,df=n-1,lower.tail=TRUE)}
        }
    }
    
    curve(dnorm(x, mu0, sigma/sqrt(n)), mu0 - 4 * sigma/sqrt(n), mu0 + 4 * sigma/sqrt(n), 
        main = paste("Sampling Distribution Under the Null Hypothesis\nType I error rate from simulation = ", 
            length(p[p < alpha])/k, sep = ""), xlab = "x", ylab = "Pr(x)", col = "red", 
        xlim = c(mu0 - 4 * sigma/sqrt(n), mu0 + 4 * sigma/sqrt(n)), ylim = c(0, 
            dnorm(mu0, mu0, sigma/sqrt(n))))
    abline(h = 0)
    
    if (alternative == "less") {
        polygon(cbind(c(mu0 - 4 * sigma/sqrt(n), seq(from = mu0 - 4 * sigma/sqrt(n), 
            to = mu0 - qnorm(1 - alpha) * sigma/sqrt(n), length.out = 100), 
            mu0 - qnorm(1 - alpha) * sigma/sqrt(n))), c(0, dnorm(seq(from = mu0 - 
            4 * sigma/sqrt(n), to = mu0 - qnorm(1 - alpha) * sigma/sqrt(n), 
            length.out = 100), mean = mu0, sd = sigma/sqrt(n)), 0), border = "black", 
            col = "grey")
        q <- pnorm(mu0 - qnorm(1 - alpha) * sigma/sqrt(n), mean = mu0, sd = sigma/sqrt(n)) - 
            pnorm(mu0 - 4 * sigma/sqrt(n), mean = mu0, sd = sigma/sqrt(n))
    }
    if (alternative == "greater") {
        polygon(cbind(c(mu0 + qnorm(1 - alpha) * sigma/sqrt(n), seq(from = mu0 + 
            qnorm(1 - alpha) * sigma/sqrt(n), to = mu0 + 4 * sigma/sqrt(n), 
            length.out = 100), mu0 + 4 * sigma/sqrt(n))), c(0, dnorm(seq(from = mu0 + 
            qnorm(1 - alpha) * sigma/sqrt(n), to = mu0 + 4 * sigma/sqrt(n), 
            length.out = 100), mean = mu0, sd = sigma/sqrt(n)), 0), border = "black", 
            col = "grey")
        q <- pnorm(mu0 + 4 * sigma/sqrt(n), mean = mu0, sd = sigma/sqrt(n)) - 
            pnorm(mu0 + qnorm(1 - alpha) * sigma/sqrt(n), mean = mu0, sd = sigma/sqrt(n))
    }
    if (alternative == "two.tailed") {
        polygon(cbind(c(mu0 - 4 * sigma/sqrt(n), seq(from = mu0 - 4 * sigma/sqrt(n), 
            to = mu0 - qnorm(1 - alpha/2) * sigma/sqrt(n), length.out = 100), 
            mu0 - qnorm(1 - alpha/2) * sigma/sqrt(n))), c(0, dnorm(seq(from = mu0 - 
            4 * sigma/sqrt(n), to = mu0 - qnorm(1 - alpha/2) * sigma/sqrt(n), 
            length.out = 100), mean = mu0, sd = sigma/sqrt(n)), 0), border = "black", 
            col = "grey")
        polygon(cbind(c(mu0 + qnorm(1 - alpha/2) * sigma/sqrt(n), seq(from = mu0 + 
            qnorm(1 - alpha/2) * sigma/sqrt(n), to = mu0 + 4 * sigma/sqrt(n), 
            length.out = 100), mu0 + 4 * sigma/sqrt(n))), c(0, dnorm(seq(from = mu0 + 
            qnorm(1 - alpha/2) * sigma/sqrt(n), to = mu0 + 4 * sigma/sqrt(n), 
            length.out = 100), mean = mu0, sd = sigma/sqrt(n)), 0), border = "black", 
            col = "grey")
        q <- pnorm(mu0 - qnorm(1 - alpha/2) * sigma/sqrt(n), mean = mu0, sd = sigma/sqrt(n)) - 
            pnorm(mu0 - 4 * sigma/sqrt(n), mean = mu0, sd = sigma/sqrt(n)) + 
            pnorm(mu0 + 4 * sigma/sqrt(n), mean = mu0, sd = sigma/sqrt(n)) - 
            pnorm(mu0 + qnorm(1 - alpha/2) * sigma/sqrt(n), mean = mu0, sd = sigma/sqrt(n))
    }
    # print(round(q,digits=3)) # this prints area in the shaded portion(s) of
    # the curve
    return(length(p[p < alpha])/k)  # returns the proportion of simulations where p < alpha
}
```
I'm not getting any output here...shouldn't curve at least be showing me a graph? I'm getting nothing. oH wait nevermind, all of that was just setting up the function I think

```{r}
eI <- typeI(mu0 = -3, sigma = 2, n = 5000, alternative = "greater", alpha = 0.05)
eI <- typeI(mu0 = 5, sigma = 2, n = 1000, alternative = "less", alpha = 0.01)
```
I'm still getting nothing, it's saying incomplete expression (jk I got it, not sure why it worked the second time)

#Challenge
From the two coding chunks above it looks like Type I error increases as "n" increases but I will check.
```{r}
eI <- typeI(mu0 = 5, sigma = 2, n = 1000, alternative = "less", alpha = 0.01)
eI <- typeI(mu0 = 5, sigma = 2, n = 2000, alternative = "less", alpha = 0.01)
eI <- typeI(mu0 = 5, sigma = 2, n = 5000, alternative = "less", alpha = 0.01)
eI <- typeI(mu0 = 5, sigma = 2, n = 10000, alternative = "less", alpha = 0.01)
```
Well the gray area doesn't seem to be changing but the x-axis is changing so I'm not too sure what conclusions can be drawn from this.

Changing sigma:
```{r}
eI <- typeI(mu0 = 5, sigma = 2, n = 1000, alternative = "less", alpha = 0.01)
eI <- typeI(mu0 = 5, sigma = 5, n = 1000, alternative = "less", alpha = 0.01)
eI <- typeI(mu0 = 5, sigma = 100, n = 1000, alternative = "less", alpha = 0.01)
```

So I guess the only thing that should change the ype I error rate would be alpha.
```{r}
eI <- typeI(mu0 = 5, sigma = 2, n = 1000, alternative = "less", alpha = 0.01)
eI <- typeI(mu0 = 5, sigma = 2, n = 1000, alternative = "less", alpha = 0.05)
eI <- typeI(mu0 = 5, sigma = 2, n = 1000, alternative = "less", alpha = 0.10) #yes that is definitely true
```

**Bonferroni correction** is a way to correct for multiple testing problems

-when you're doing "k" independent hypothesis tests with a significance level of alpha, adjust the alpha level used to interpret statistical significance to:

alpha(B)=alpha/k

-example: if we run 10 hypothesis tests (k=10), the adjusted alpha level would be 0.05/10=0.005

-this is also called "limiting the family-wise error rate to level alpha"

-this correction is considered very conservative, so there are other options
```{r Bonferroni correction}
alpha <- 0.05
pvals <- c(1e-04, 0.003, 0.005, 0.01, 0.02, 0.04, 0.045, 0.11, 0.18, 0.23)
sig <- pvals <= alpha/length(pvals)
sig  # first 3 values are less than the adjusted alpha
```

**Benjamini & Hochberg correction** is less conservative than Bonferroni

-goal of this one is to control the FDR (in this case incorrect rejections of the null) 
```{r Benjamini and Hochberg}
library(ggplot2)
alpha <- 0.05
psig <- NULL
pvals <- c(1e-04, 0.003, 0.005, 0.01, 0.02, 0.04, 0.045, 0.11, 0.18, 0.27)
for (i in 1:length(pvals)) {
    psig[i] <- alpha * i/length(pvals)
}
d <- data.frame(cbind(rank = c(1:10), pvals, psig))
p <- ggplot(data = d, aes(x = rank, y = pvals)) + geom_point() + geom_line(aes(x = rank, 
    y = psig))
p
sig <- pvals <= psig  # vector of significant pvalues
sig
```
We can also change the p values themselves instead of the alpha levels
```{r adjusting p values}
sig <- p.adjust(pvals, method = "bonferroni") <= 0.05
sig
sig <- p.adjust(pvals, method = "BH") <= 0.05
sig
```

#Type II Error
-reducing the alpha reduces the chances of a Type I error but it DIRECTLY increases the chance of a Type II error (incorrectly accepting the null)

-we can't predict beta because we don't know where the distribution is truly centered; we can simulate what beta should look like under different Ha's, sample sizes, and alphas

-we're doing the same type of thing function-wise as typeI(), where we specify mu(A) so we get different alternative hypotheses
```{r type II function}
typeII <- function(mu0, muA, sigma, n, alternative = "two.tailed", alpha = 0.05, 
    k = 1000) {
    p <- rep(NA, k)  # sets up a vector of empty p values
    for (i in 1:k) {
        x <- rnorm(n = n, mean = muA, sd = sigma)  # draw from Ha
        m <- mean(x)
        s <- sd(x)
        z <- (m - mu0)/(s/sqrt(n))  # calculates the Z statistic for the sample drawn from Ha relative to the null distribution
        if (alternative == "less") {
            p[i] <- pnorm(z, lower.tail = TRUE)  # calculates the associated p value
            hyp <- "muA < mu0"
        }
        if (alternative == "greater") {
            p[i] <- pnorm(z, lower.tail = FALSE)
            hyp <- "muA > mu0"
        }
        if (alternative == "two.tailed") {
            if (z > 0) {
                p[i] <- 2 * pnorm(z, lower.tail = FALSE)
            }
            if (z < 0) {
                p[i] <- 2 * pnorm(z, lower.tail = TRUE)
            }
            hyp <- "muA ≠ mu0"
        }
    }
    
    curve(dnorm(x, mu0, sigma/sqrt(n)), mu0 - 4 * sigma/sqrt(n), mu0 + 4 * sigma/sqrt(n), 
        main = paste("Sampling Distributions Under the Null (red)\nand Alternative Hypotheses (blue)\nType II error rate from simulation = ", 
            length(p[p >= alpha])/k, sep = ""), xlab = "x", ylab = "Pr(x)", 
        col = "red", xlim = c(min(c(mu0 - 4 * sigma/sqrt(n), muA - 4 * sigma/sqrt(n))), 
            max(c(mu0 + 4 * sigma/sqrt(n), muA + 4 * sigma/sqrt(n)))), ylim = c(0, 
            max(c(dnorm(mu0, mu0, sigma/sqrt(n))), dnorm(muA, muA, sigma/sqrt(n)))))
    
    curve(dnorm(x, muA, sigma/sqrt(n)), muA - 4 * sigma/sqrt(n), muA + 4 * sigma/sqrt(n), 
        col = "blue", add = TRUE)
    abline(h = 0)
    
    if (alternative == "less") {
        polygon(cbind(c(mu0 - qnorm(1 - alpha) * sigma/sqrt(n), seq(from = mu0 - 
            qnorm(1 - alpha) * sigma/sqrt(n), to = muA + 4 * sigma/sqrt(n), 
            length.out = 100), muA + 4 * sigma/sqrt(n))), c(0, dnorm(seq(mu0 - 
            qnorm(1 - alpha) * sigma/sqrt(n), to = muA + 4 * sigma/sqrt(n), 
            length.out = 100), mean = muA, sd = sigma/sqrt(n)), 0), border = "black", 
            col = "grey")
        abline(v = mu0 - qnorm(1 - alpha) * sigma/sqrt(n), col = "black", lty = 3, 
            lwd = 2)
    }
    
    if (alternative == "greater") {
        polygon(cbind(c(muA - 4 * sigma/sqrt(n), seq(from = muA - 4 * sigma/sqrt(n), 
            to = mu0 + qnorm(1 - alpha) * sigma/sqrt(n), length.out = 100), 
            mu0 + qnorm(1 - alpha) * sigma/sqrt(n))), c(0, dnorm(seq(from = muA - 
            4 * sigma/sqrt(n), to = mu0 + qnorm(1 - alpha) * sigma/sqrt(n), 
            length.out = 100), mean = muA, sd = sigma/sqrt(n)), 0), border = "black", 
            col = "grey")
        abline(v = mu0 + qnorm(1 - alpha) * sigma/sqrt(n), col = "black", lty = 3, 
            lwd = 2)
    }
    
    if (alternative == "two.tailed") {
        abline(v = mu0 - qnorm(1 - alpha/2) * sigma/sqrt(n), col = "black", 
            lty = 3, lwd = 2)
        abline(v = mu0 + qnorm(1 - alpha/2) * sigma/sqrt(n), col = "black", 
            lty = 3, lwd = 2)
        
        if (z > 0) {
            # greater
            polygon(cbind(c(muA - 4 * sigma/sqrt(n), seq(from = muA - 4 * sigma/sqrt(n), 
                to = mu0 + qnorm(1 - alpha/2) * sigma/sqrt(n), length.out = 100), 
                mu0 + qnorm(1 - alpha/2) * sigma/sqrt(n))), c(0, dnorm(seq(from = muA - 
                4 * sigma/sqrt(n), to = mu0 + qnorm(1 - alpha/2) * sigma/sqrt(n), 
                length.out = 100), mean = muA, sd = sigma/sqrt(n)), 0), border = "black", 
                col = "grey")
        }
        
        # less
        if (z < 0) {
            polygon(cbind(c(mu0 - qnorm(1 - alpha/2) * sigma/sqrt(n), seq(from = mu0 - 
                qnorm(1 - alpha/2) * sigma/sqrt(n), to = muA + 4 * sigma/sqrt(n), 
                length.out = 100), muA + 4 * sigma/sqrt(n))), c(0, dnorm(seq(mu0 - 
                qnorm(1 - alpha/2) * sigma/sqrt(n), to = muA + 4 * sigma/sqrt(n), 
                length.out = 100), mean = muA, sd = sigma/sqrt(n)), 0), border = "black", 
                col = "grey")
        }
    }
    
    return(length(p[p >= alpha])/k)
}
```
#Challenge
```{r Changing muA values}
eII <- typeII(mu0 = 2, muA = 4, sigma = 3, n = 6, alternative = "greater")
eII <- typeII(mu0 = 2, muA = 0.2, sigma = 3, n = 6, alternative = "greater")
eII <- typeII(mu0 = 2, muA = 20, sigma = 3, n = 6, alternative = "greater")
```

It looks like a larger muA decreases the likelihood of a Type II error? There's no gray area in the third graph because there's no overlap on the two curves

```{r Changing sigma}
eII <- typeII(mu0 = 2, muA = 4, sigma = 3, n = 6, alternative = "greater")
eII <- typeII(mu0 = 2, muA = 4, sigma = 1, n = 6, alternative = "greater")
eII <- typeII(mu0 = 2, muA = 4, sigma = 30, n = 6, alternative = "greater")
```

Smaller sigma results in smaller gray area, meaning decreased chance of Type II Error
```{r Changing n}
eII <- typeII(mu0 = 2, muA = 4, sigma = 3, n = 6, alternative = "greater")
eII <- typeII(mu0 = 2, muA = 4, sigma = 3, n = 2, alternative = "greater")
eII <- typeII(mu0 = 2, muA = 4, sigma = 3, n = 60, alternative = "greater")
```

Increasing n (sample size) decreases the chances of a Type II error...does this mean that beta is smaller?
```{r one and two tailed tests}
eII <- typeII(mu0 = 2, muA = 4, sigma = 3, n = 6, alternative = "greater")
eII <- typeII(mu0 = 2, muA = 4, sigma = 3, n = 6, alternative = "two.tailed")
eII <- typeII(mu0 = 2, muA = 4, sigma = 3, n = 6, alternative = "less")
```

I don't really understand what this means, "less" results in the whole alternative hypothesis curve being colored gray...does this mean that it's incredibly likely for a Type II Error? Or is it because in this case it would never be less because mu0 is less than the set muA?

Two tailed has a larger gray area than the original one-tailed test...this makes sense in my head but I'm not sure I could explain why. Does it have something to do with looking at 2.5% of data on either end versus just one? (Like you don't know if muA or whatever is larger or smaller than mu0)

**power** is the proability of correctly rejecting a null hypothesis; can be represented as 1-beta

-power values of 0.8 or higher are high

**effect size** is a standardized difference between the means of the groups being compared; indicator of the strength of a phenomenon

-divide the difference between the means by the standard deviation

-values of 0.2 or less are considered low and values of 0.8 and above are considered high
```{r manipulating power and effect size}
library(ggplot2)
library(manipulate)
power.plot <- function(sigma, muA, mu0, n, alpha, alternative = "two.tailed") {
    pow <- 0
    z <- (muA - mu0)/(sigma/sqrt(n))
    g <- ggplot(data.frame(mu = c(min(mu0 - 4 * sigma/sqrt(n), muA - 4 * sigma/sqrt(n)), 
        max(mu0 + 4 * sigma/sqrt(n), muA + 4 * sigma/sqrt(n)))), aes(x = mu)) + 
        ggtitle("Explore Power for Z Test")
    g <- g + ylim(c(0, max(dnorm(mu0, mu0, sigma/sqrt(n)) + 0.1, dnorm(muA, 
        muA, sigma/sqrt(n)) + 0.1)))
    g <- g + stat_function(fun = dnorm, geom = "line", args = list(mean = mu0, 
        sd = sigma/sqrt(n)), size = 1, col = "red", show.legend = TRUE)
    g <- g + stat_function(fun = dnorm, geom = "line", args = list(mean = muA, 
        sd = sigma/sqrt(n)), size = 1, col = "blue", show.legend = TRUE)
    
    if (alternative == "greater") {
        if (z > 0) {
            xcrit = mu0 + qnorm(1 - alpha) * sigma/sqrt(n)
            g <- g + geom_segment(x = xcrit, y = 0, xend = xcrit, yend = max(dnorm(mu0, 
                mu0, sigma/sqrt(n)) + 0.025, dnorm(muA, muA, sigma/sqrt(n)) + 
                0.025), size = 0.5, linetype = 3)
            g <- g + geom_polygon(data = data.frame(cbind(x = c(xcrit, seq(from = xcrit, 
                to = muA + 4 * sigma/sqrt(n), length.out = 100), muA + 4 * sigma/sqrt(n)), 
                y = c(0, dnorm(seq(from = xcrit, to = muA + 4 * sigma/sqrt(n), 
                  length.out = 100), mean = muA, sd = sigma/sqrt(n)), 0))), 
                aes(x = x, y = y), fill = "blue", alpha = 0.5)
            pow <- pnorm(muA + 4 * sigma/sqrt(n), muA, sigma/sqrt(n)) - pnorm(xcrit, 
                muA, sigma/sqrt(n))
        }
    }
    if (alternative == "less") {
        if (z < 0) {
            xcrit = mu0 - qnorm(1 - alpha) * sigma/sqrt(n)
            g <- g + geom_segment(x = xcrit, y = 0, xend = xcrit, yend = max(dnorm(mu0, 
                mu0, sigma/sqrt(n)) + 0.025, dnorm(muA, muA, sigma/sqrt(n)) + 
                0.025), size = 0.5, linetype = 3)
            g <- g + geom_polygon(data = data.frame(cbind(x = c(muA - 4 * sigma/sqrt(n), 
                seq(from = muA - 4 * sigma/sqrt(n), to = xcrit, length.out = 100), 
                xcrit), y = c(0, dnorm(seq(from = muA - 4 * sigma/sqrt(n), to = xcrit, 
                length.out = 100), mean = muA, sd = sigma/sqrt(n)), 0))), aes(x = x, 
                y = y), fill = "blue", alpha = 0.5)
            pow <- pnorm(xcrit, muA, sigma/sqrt(n)) - pnorm(muA - 4 * sigma/sqrt(n), 
                muA, sigma/sqrt(n))
        }
    }
    if (alternative == "two.tailed") {
        if (z > 0) {
            xcrit = mu0 + qnorm(1 - alpha/2) * sigma/sqrt(n)
            g <- g + geom_segment(x = xcrit, y = 0, xend = xcrit, yend = max(dnorm(mu0, 
                mu0, sigma/sqrt(n)) + 0.025, dnorm(muA, muA, sigma/sqrt(n)) + 
                0.025), size = 0.5, linetype = 3)
            g <- g + geom_polygon(data = data.frame(cbind(x = c(xcrit, seq(from = xcrit, 
                to = muA + 4 * sigma/sqrt(n), length.out = 100), muA + 4 * sigma/sqrt(n)), 
                y = c(0, dnorm(seq(from = xcrit, to = muA + 4 * sigma/sqrt(n), 
                  length.out = 100), mean = muA, sd = sigma/sqrt(n)), 0))), 
                aes(x = x, y = y), fill = "blue", alpha = 0.5)
            pow <- pnorm(muA + 4 * sigma/sqrt(n), muA, sigma/sqrt(n)) - pnorm(xcrit, 
                muA, sigma/sqrt(n))
        }
        if (z < 0) {
            xcrit = mu0 - qnorm(1 - alpha/2) * sigma/sqrt(n)
            g <- g + geom_segment(x = xcrit, y = 0, xend = xcrit, yend = max(dnorm(mu0, 
                mu0, sigma/sqrt(n)) + 0.025, dnorm(muA, muA, sigma/sqrt(n)) + 
                0.025), size = 0.5, linetype = 3)
            g <- g + geom_polygon(data = data.frame(cbind(x = c(muA - 4 * sigma/sqrt(n), 
                seq(from = muA - 4 * sigma/sqrt(n), to = xcrit, length.out = 100), 
                xcrit), y = c(0, dnorm(seq(from = muA - 4 * sigma/sqrt(n), to = xcrit, 
                length.out = 100), mean = muA, sd = sigma/sqrt(n)), 0))), aes(x = x, 
                y = y), fill = "blue", alpha = 0.5)
            pow <- pnorm(xcrit, muA, sigma/sqrt(n)) - pnorm(muA - 4 * sigma/sqrt(n), 
                muA, sigma/sqrt(n))
        }
    }
    g <- g + annotate("text", x = max(mu0, muA) + 2 * sigma/sqrt(n), y = max(dnorm(mu0, 
        mu0, sigma/sqrt(n)) + 0.075, dnorm(muA, muA, sigma/sqrt(n)) + 0.075), 
        label = paste("Effect Size = ", round((muA - mu0)/sigma, digits = 3), 
            "\nPower = ", round(pow, digits = 3), sep = ""))
    g <- g + annotate("text", x = min(mu0, muA) - 2 * sigma/sqrt(n), y = max(dnorm(mu0, 
        mu0, sigma/sqrt(n)) + 0.075, dnorm(muA, muA, sigma/sqrt(n)) + 0.075), 
        label = "Red = mu0\nBlue = muA")
    g
}
manipulate(power.plot(sigma, muA, mu0, n, alpha, alternative), sigma = slider(1, 
    10, step = 1, initial = 4), muA = slider(-10, 10, step = 1, initial = 2), 
    mu0 = slider(-10, 10, step = 1, initial = 0), n = slider(1, 50, step = 1, 
        initial = 16), alpha = slider(0.01, 0.1, step = 0.01, initial = 0.05), 
    alternative = picker("two.tailed", "greater", "less"))
```

power.t.test() gives us calculations based on a t distribution (what we will normally be using)
#Challenge
```{r}
library(ggplot2)
library(manipulate)
power.test <- function(mu0, muA, sigma, alpha = 0.05, type, alternative) {
    p <- 0
    for (i in 2:200) {
        x <- power.t.test(n = i, delta = abs(muA - mu0), sd = sigma, sig.level = alpha, 
            power = NULL, type = type, alternative = alternative)
        p <- c(p, x$power)
    }
    d <- data.frame(cbind(1:200, p, 1 - p))
    critn <- 0
    for (i in 1:199) {
        if (p[i] < 0.8 && p[i + 1] >= 0.8) {
            critn <- i + 1
        } else {
            critn <- critn
        }
    }
    names(d) <- c("n", "power", "beta")
    g <- ggplot(data = d) + xlab("sample size n") + ylab("Type II Error Rate, Beta  (Red)\nand\nPower, 1-Beta (Blue)") + 
        ggtitle("Power for T Tests\n(assuming equal n and variance across the two groups)") + 
        ylim(0, 1) + geom_point(aes(x = n, y = power), colour = "blue", alpha = 1/2) + 
        geom_line(aes(x = n, y = power), colour = "blue", alpha = 1/2) + geom_line(aes(x = n, 
        y = 0.8), colour = "red", lty = 3) + geom_point(aes(x = n, y = beta), 
        colour = "red", alpha = 1/2) + geom_line(aes(x = n, y = beta), colour = "red", 
        alpha = 1/2) + geom_linerange(aes(x = critn, ymin = 0, ymax = 0.8), 
        colour = "blue", alpha = 1/4) + annotate("text", x = 150, y = 0.5, label = paste("Effect Size = ", 
        round(abs(mu0 - muA)/sigma, digits = 3), "\nCritical n = ", critn, sep = ""))
    print(g)
}
manipulate(power.test(mu0, muA, sigma, alpha, type, alternative), mu0 = slider(-10, 
    10, initial = 3, step = 1), muA = slider(-10, 10, initial = 0, step = 1), 
    sigma = slider(1, 10, initial = 3, step = 1), alpha = slider(0.01, 0.1, 
        initial = 0.05, step = 0.01), alternative = picker("two.sided", "one.sided"), 
    type = picker("two.sample", "one.sample", "paired"))
```

