---
title: "Swirl-GGplot 1"
author: "Dani Antos"
date: "October 1, 2017"
output: html_document
---

ggplot2 is the newest plotting system in R. "Aesthetics" allow us to customize attributes like color, shape, and size, and geometric objects (geoms) like points, lines, and bars. 

ggplot2 is the best of "base" and "lattice"- allows for multipanel plots but allows you to add stuff. 2 workhouse functions: **(1)** qplot, easier of the two; and **(2)** ggplot, more flexible for customisations

```{r}
library(ggplot2)
str(mpg)
qplot(displ, hwy, data = mpg)
qplot(displ, hwy, data = mpg, color=drv) #color codes points
qplot(displ, hwy, data = mpg, color=drv, geom= c("point", "smooth")) #gives trendlines and 95% CI's
qplot(y=hwy, data= mpg, color=drv)
qplot(drv, hwy, data= mpg, geom= "boxplot")
qplot(drv, hwy, data= mpg, geom= "boxplot", color= manufacturer)
```

Moving on to histograms:
```{r}
qplot(hwy, data= mpg, fill=drv)
qplot(displ, hwy, data= mpg, facets= . ~drv) #makes scatterplot with 3 facets or sections
qplot(hwy, data= mpg, facets= drv ~ ., binwidth= 2) #faceted barplot
```

