---
title: "Swirl-Getting and Cleaning Data"
author: "Dani Antos"
date: "September 19, 2017"
output: html_document
---

from the previous lesson, summarize() is the most powerful manipulations to apply to grouped data
```{r creating cran}
library(dplyr)
cran <- tbl_df(mydf)
rm("mydf")
cran
```

To group cran by the "package" variable:
```{r summarize function}
by_package <- group_by(cran, package, add=FALSE)
by_package
summarize(by_package, mean(size)) #returns the mean size of each package instead of a single value
pack_sum <- summarize(by_package,
                      count = n(),
                      unique = n_distinct(ip_id),
                      countries = n_distinct(country),
                      avg_bytes = mean(size)) #I didn't really do much for this, just copied and pasted
pack_sum
```

To isolate a certain percentage of the data, we can use percentiles and quantile functions
```{r quantiles}
quantile(pack_sum$count, probs = 0.99)
top_counts <- filter(pack_sum, count>679) #selects rows that have at least 679 counts
top_counts
View(top_counts) #capital V, shows all rows
top_counts_sorted <- arrange(top_counts, desc(count)) #top counts sorted from greatest to least
View(top_counts_sorted)
```

Now we're doing quantiles of the unique column 
```{r}
quantile(pack_sum$unique, probs= 0.99)
top_unique <- filter(pack_sum, unique>465)
View(top_unique)
top_unique_sorted <- arrange(top_unique, desc(unique))
View(top_unique_sorted)
```

#Chaining
```{r}
by_package <- group_by(cran, package)
pack_sum <- summarize(by_package,
                      count = n(),
                      unique = n_distinct(ip_id),
                      countries = n_distinct(country),
                      avg_bytes = mean(size))
top_countries <- filter(pack_sum, countries > 60)
result1 <- arrange(top_countries, desc(countries), avg_bytes)
print(result)
```
Sorted primarily by country but used average bytes as a tie-breaker

```{r}

result2 <-
  arrange(
    filter(
      summarize(
        group_by(cran,
                 package
        ),
        count = n(),
        unique = n_distinct(ip_id),
        countries = n_distinct(country),
        avg_bytes = mean(size)
      ),
      countries > 60
    ),
    desc(countries),
    avg_bytes
  )

print(result2)

```

#Chaining
```{r}
result3 <-
  cran %>%
  group_by(package) %>%
  summarize(count = n(),
            unique = n_distinct(ip_id),
            countries = n_distinct(country),
            avg_bytes = mean(size)
  ) %>%
  filter(countries > 60) %>%
  arrange(desc(countries), avg_bytes)
print(result3)
View(result3) #view dataset, don't forget the capital V
```
The last three code chunks all have the same output, but the last one is more convenient and concise.

```{r}
cran %>%
  select(ip_id, country, package, size) %>%
  mutate(size_mb = size / 2^20) %>%
  filter(size_mb <= 0.5) %>%
  arrange(desc(size_mb)) %>%
  print #this is the whole script, in the tutorial I just took it one line at a time
```

