---
title: "Swirl- GGplot 2"
author: "Dani Antos"
date: "October 1, 2017"
output: html_document
---

ggplot part 2, focusing on the fundamental components of the package to be better able to customise plots. We're also going over the second workhorse function **ggplot**.

There are 7 main components of ggplot2 plots:
-data frame: holds the data to be plotted
-aesthetic mapping: how data are mapped to color, size, etc.
-geoms: what we see in the plots; lines, points, etc. 
-facets: panels used in conditional plots
-stats: statistical transformations like binning, quantiles, etc.
-scales: cording that the aesthetic map uses
-coordinate system: how the plots are depicted
the last three are taken care of automatically in qplot

```{r}
qplot(displ, hwy, data= mpg, geom= c("point", "smooth"), facets= .~drv) #recap on how to use qplot
g <- ggplot(mpg, aes(displ,hwy))
summary(g)
g+geom_point()
g+geom_point()+geom_smooth() #adds trendline and confidence band
g+geom_point()+geom_smooth(method="lm") #trendline of linear model
g+geom_point()+geom_smooth(method="lm")+facet_grid( .~drv) #three different panels
g+geom_point()+geom_smooth(method="lm")+facet_grid( .~drv)+ggtitle("Swirl Rules!") #adds title
```

Now we're going to practice customising
```{r}
g+geom_point(color="pink", size=4, alpha=1/2) #alpha sets the transparency
g + geom_point(aes(color = drv), size = 4, alpha = 1/2) #we have to use aes because the color is data dependent
g + geom_point(aes(color = drv)) + labs(title="Swirl Rules!") + labs(x="Displacement", y="Hwy Mileage") #way to add x and y axis titles
g+geom_point(aes(color=drv), size=2, alpha=1/2)+geom_smooth(size=4, linetype=3, method="lm", se=FALSE) #method specifices linear regression, linetype signifies dashed
g+geom_point(aes(color=drv))+theme_bw(base_family="Times")
plot(myx,myy, type="l", ylim=c(-3,3)) #base function plotg<-
```

```{r}
g<- ggplot(testdat, aes(myx,myy))
g+geom_line() #displays outlier point, can't see a lot of the other data
g+geom_line()+ylim(-3,3) #adds limit, ggplot ignores outlier
g+geom_line()+coord_cartesian(ylim=c(-3,3)) #looks more like the basic plot function
```

