---
title: "Swirl-Getting and Cleaning Data"
author: "Dani Antos"
date: "September 19, 2017"
output: html_document
---


```{r}
library(dplyr)
mydf<-read.csv(path2csv,stringsAsFactors = FALSE)
dim(mydf)
head(mydf)
packageVersion("dplyr")
```

First step is to load data into a data frame tbl
```{r}
cran <- tbl_df(mydf)
rm("mydf") #used to remove the original dataframe
cran #used to show the data in the console
```
this allows data to be more informative and compact than an original data frame

The five "verbs" that cover most fundamental data manipulation tasks: select(), filter(), arrange(), mutate(),
summarize()

in large datasets, you usually only want some of the variables
```{r}
select(cran, ip_id, package, country) #only shows those three columns chosen, using select function is specific to columns, and they are returned in the order specified in the code
```

using select, we can also specify a sequence of columns using a colon (similar to numbers in vectors)
```{r}
select(cran, r_arch:country)
select(cran, country:r_arch) #gives column sequence in reverse order
cran #to show entire dataset again
select(cran, -time) #used to get rid of a column, in this case "time"
```

Omitting columns and grabbing sequences of columns can be combined to omit sequences of columns
```{r}
select(cran, -(x:size)) #make sure to use X not x
```

Now we're filtering row instead of columns, using filter()
```{r}
filter(cran, package == "swirl") #filter recognizes package as a column of cran
filter(cran, r_version == "3.1.1", country == "US") #rows corresponding to downloads from users in the us running that specific version of R
filter(cran, r_version <= "3.0.2", country == "IN") #R version will be less than or equal to 3.0.2
filter(cran, country == "US" | country == "IN")
 filter(cran, size >100500, r_os=="linux-gnu") #doesn't need quotes because it is a number
```

NA values can be detected with is.na()
```{r}
is.na(c(3,5,NA,10))
!is.na(c(3,5,NA,10)) #adding '!' flips all of the values, ie. true becomes false
filter(cran, !is.na(r_version)) 
```

Arrange() allows you to order the rows of a dataset according to the values of a particular variable
```{r}
cran2<-select(cran, size:ip_id)
arrange(cran2, ip_id) #ip_id column arranged in ascending order
arrange(cran2, desc(ip_id)) #arrange it in descending order
```

we can also arrange by multiple variables
```{r}
arrange(cran2, package, ip_id)
arrange(cran2, country, desc(r_version), ip_id)
```

mutate() function
```{r}
cran3 <- select(cran, ip_id, package, size)
cran3
mutate(cran3, size_mb = size / 2^20) #added a column that displayed the size in mb instead of bytes
mutate(cran3, size_mb=size/2^20, size_gb=size_mb/2^10)
mutate(cran3, correct_size = size + 1000)
```

summarize() collapses the dataset into one row
```{r}
summarize(cran, avg_bytes= mean(size))

```

